---
title: "DS250 - Introduction to Data Science"
author: "Kevin Niemann"
date: "November 3, 2015"
output:  html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
rm( list = ls())  # Clear environment
```

# Assignment 5 - Core concepts in Statistics

This assignment covers core concepts in data and descriptive and inferential statistics. Statistical analysis is one of the primary tools of a Data Scientist used in understanding data, testing validity, and making predictions.

### The objectives of this assignment are:
* Understand the types of data
* Review core concepts in descriptive statistics
* High level review of probability distributions
* Practice making predictions using regression and probability

***
#### Question 1 - Types of Data
**a**\. Describe some of the differences between qualitative and quantitative data:

````
Quantitative data can be measured numerically, may be considered objective. Qualitative is a description of something. This could be subjective. 

````

**b**\. Describe some of the differences between an interval and a ratio:

````
The main difference is that an interval of zero doesn't mean zero of that thing, versus a ratio of zero does.
````

**c**\. From the image below, identify the statistical data types in the departure table:(http://canvas.uw.edu/files/33441437/download?download_frd=1)
Quantitative: destination, flight, airline, term gate, Status, Equip
Qualitative: On-time rating, Departure
***
#### Question 2 - Distributions
**a**\. Central Tendency: Briefly describe the following measures:

*    Mode: the value with the highest occurance.
*    Median: the number that is halfway through a set.
*    Mean: the sum of a set of numbers divided by the number of numbers     (average).

````

````

**b**\. Dispersion: Briefly describe the following measures:

*    Range: the difference between the highest and lowest numbers in a set.
*    Variance: the average of the squared distances from the mean.
*    Standard deviation: The square root of the variance.

````

````

**c**\. Using the following sequence, create a plot of the normal distribution with a mean = 0 and standard deviation = 1.

```{r,echo=TRUE}
x <- seq(-5,3, 0.01)
plot(x=x, y=dnorm(x, mean=0, sd=1), type='l')

````

***
#### Question 3 - Probabilities
**a**\. Using the heart patient data in the 'processed.cleveland.data.txt' file: (link below)   
(https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/heart-disease.names)  
(https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data)  

- separate the population into male and female groups  
- calculate the mean and standard deviation for male cholesterol values  
- create a histogram showing the distribution of male cholesterol values 

```{r,echo=TRUE}
heart <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data",sep=",",header=FALSE)
colnames(heart) <- c("age","sex","cp","trestbps","chol","fbs","restecg","thalach","exang","oldpeak","slope","ca","thal","num")
mheart <- subset(heart, heart$sex == 1)
fheart <- subset(heart, heart$sex == 0)
mean(mheart$chol)
hist(mheart$chol, probability = TRUE)

````

**b**\. Using **rnorm()**, create a normal distribution and compare the histogram of the male cholesterol against the simulated normal distribution. Are they the same? *(can also use dnorm())*

```
mcholmean <- mean(mheart$chol)
mcholsd   <- sd(mheart$chol)
x <- 126:353
y <- rnorm(n=x, mean = mcholmean, sd = mcholsd)
hist(y,probability = TRUE)

Based on the histogram comparison, the data looks like a normal distribution.
````

**c**\. Using **pnorm()**, determine the probability that a randomly chosen adult male has a cholesterol level of over 260.

````
1 - pnorm(q = 260, mean = mcholmean, sd = mcholsd)
.31
````

***
#### Question 4 - Comparison
**a**\. Load sample data file from: (http://www.openintro.org/stat/data/cdc.R)  
*(use "source(file)" function to load external R data.)*

````
source("http://www.openintro.org/stat/data/cdc.R")
````

**b**\. Create a histogram plot of the weight values:

````
hist(cdc$weight)
````

**c**\. Create a new data frame (*young*) from the cdc data file containing individuals between the ages of (18 to 30).

````
young <- subset(cdc, age >= 18 & age <= 30)
````

**d**\. Use **sample()**, to collect 50 random sample values from the (young) data frame.

````
sample <- young[sample(nrow(young), 50, replace = FALSE, prob = NULL),]
````

**e**\. Describe the distribution of this sample?  
How does it compare to the distribution of the population from which it was sampled from?

````
cdcNum <- sapply(cdc, is.numeric)
colMeans(cdc[, cdcNum])

sampleNum <- sapply(sample, is.numeric)
colMeans(sample[, sampleNum])

Sample is taller, weighs less, smokes less.

````

**f**\. Using **sample()**, create three separate sample collections for (10, 50, 100) samples and plot the histogram of these three data sets. How are they different?  

*using the original cdc data, not the filtered 'young' data*

````
sample10 <- cdc[sample(nrow(cdc), 10, replace = FALSE, prob = NULL),]
hist(sample10$age)
sample50 <- cdc[sample(nrow(cdc), 50, replace = FALSE, prob = NULL),]
hist(sample50$age)
sample100 <- cdc[sample(nrow(cdc), 100, replace = FALSE, prob = NULL),]
hist(sample100$age)

````

***
#### Question 5 - Confidence Intervals  
Using the sample data from: (http://www.openintro.org/stat/data/cdc.R)  
*(use "source()" to load an external file)*


**a**\. Create a data frame with the height values from the cdc sample data.

````
height <- data.frame(cdc$height)
````

**b**\. Create a sample data frame with 60 random samples of individual height values. Based on this sample data, what it the average height for these 60 individuals?

````
sample60 <- data.frame(height=height[sample(nrow(height), 60, replace = FALSE, prob = NULL),])
mean(sample60$height)
67.16
````

**c**\. Calculate a confidence interval (upper, lower bounds) for the sample data.

````
t <- t.test(sample60$height)
t

````

**d**\. Write a short loop to create (50) separate sample sets and calculate the confidence intervals for each set. How do the confidence intervals change in the 50 samples?


````
intervals <- list()
for (i in 1:50) {
   sampleset <- sample_n(height,60,replace=FALSE)
   intervals[[i]] <- t.test(sampleset)$conf.int
}

intervals.df <- as.data.frame(do.call(rbind, intervals))

ggplot(data = intervals.df) + geom_segment(aes(x = 1:50, xend = 1:50, y = V1, yend = V2))

The confidence intervals change randomly.
````

***
#### Question 6 - Inference  
Use **"load()"** to import sample data file from: (http://www.openintro.org/stat/data/nc.RData) and complete the following:

**a**\. Review the attributes in this data set and determine which ones are categorical or numeric.

````
load(url('http://www.openintro.org/stat/data/nc.RData'))
head(nc)
Numerical: fage,mage,weeks,visits,gained,weight
Categorical: mature,premie,martial,lowbirthweight,gender,habit,whitemom

````

**b**\. Create a side-by-side boxplot of habit and weight. What does the plot highlight about the relationship between these two variables?

````
library(ggplot2)
data <- tbl_df(nc)
plot<-ggplot(data, aes(x=habit, y=weight))
plot+geom_boxplot()

Smokers kids weight less than non smoker kids at birth.
````

**c**\. Using the following statement to compare weight and habit. Is the difference between the mean of nonsmoker and that of smoker statistically significant?

````
by(nc$weight, nc$habit, mean)

g1 <- tnc %>% filter(habit == 'smoker') %>% select(weight)
g2 <- tnc %>% filter(habit == 'nonsmoker') %>% select(weight)
t.test(g2$weight,g1$weight, paired=FALSE, var.equal=FALSE)$conf

Difference is statistically different.

````

**d**\. Use the following statement to calculate the inference measure between weight and habit.

````
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")

Response variable: numerical, Explanatory variable: categorical
Difference between two means
Summary statistics:
n_nonsmoker = 873, mean_nonsmoker = 7.1443, sd_nonsmoker = 1.5187
n_smoker = 126, mean_smoker = 6.8287, sd_smoker = 1.3862
Observed difference between means (nonsmoker-smoker) = 0.3155

H0: mu_nonsmoker - mu_smoker = 0 
HA: mu_nonsmoker - mu_smoker != 0 
Standard error = 0.134 
Test statistic: Z =  2.359 
p-value =  0.0184 



````

**e**\. In the statement above, change the type argument to 'ci' to construct a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers.

````
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, alternative = "twosided", method = "theoretical")
Response variable: numerical, Explanatory variable: categorical
Difference between two means
Summary statistics:
n_nonsmoker = 873, mean_nonsmoker = 7.1443, sd_nonsmoker = 1.5187
n_smoker = 126, mean_smoker = 6.8287, sd_smoker = 1.3862
Observed difference between means (nonsmoker-smoker) = 0.3155

Standard error = 0.1338 
95 % Confidence interval = ( 0.0534 , 0.5777 )

````

***
#### Question 7 - Linear Regression  
Load the sample data file from: (http://www.openintro.org/stat/data/mlb11.RData) and complete the following:

**a**\. Use the **cor()** function to determine the numeric relationship between 'runs' and 'at_bats'.

````
load(url('http://www.openintro.org/stat/data/mlb11.RData'))

cor(x=mlb11$at_bats,y=mlb11$runs)

````

**b**\. Use the custom function **plot_ss()** *(included in RData file)* to create a plot for:

````
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)

````

**c**\. Review the mlb11 data set and select another variable that might be a good predictor of runs.

````
cor(x=mlb11$new_onbase,y=mlb11$runs)
On base (percentage of time a batter gets to a base)

````

**d**\. Create a scatterplot of the two variables and fit a linear model. Determine if there is a linear relationship?

```
plot <- ggplot(mlb11, aes(new_onbase, runs))
plot + geom_smooth(method=lm)

mlb11.lm <- lm(formula = runs ~ new_onbase, data = mlb11)
summary(mlb11.lm)
```
P value is very small so there is definitely a linear relationship.

***
#### Question 8 - Linear Regression  
Use the built-in dataset (*faithful*) and complete the following:

a) Create a linear model *lm()* and capture the standard residuals :

````
data(faithful)

faithlm <- lm(eruptions ~ waiting, data=faithful)
faithresid = rstandard(faithlm) 
summary(faithlm)

````

b) Using *qqnorm()* create a normal probability plot with *qqline()* showing the comparison:

````
qqnorm(faithresid, ylab="Residuals", xlab="Score", main="Faithful eruptions") 

qqline(faithresid) 
````


#### **End**
